Chord: A Static and Dynamic Program Analysis Framework for Java

### 2010-12-13 main/doc/building_scope.tex
migrated documentation from latex2html to hyperlatex

\Chord\ computes the scope (i.e., reachable methods) of the given
program either if property \code{chord.build.scope} is set to {\tt
  true} or if some other task (e.g., a program analysis specified via
property \code{chord.run.analyses}) demands it by calling method
\code{chord.program.Program.g()}.

@@

If property \code{chord.reuse.scope} has value {\tt true} and both
files specified by properties \code{chord.methods.file} and
\code{chord.reflect.file} exist, then \Chord\ regards those files as
specifying which methods to consider reachable and how to resolve
reflection, respectively.

@@

If property \code{chord.reuse.scope} has value {\tt false} or the
files specified by properties \code{chord.methods.file} or
\code{chord.reflect.file} do not exist, then \Chord\ computes program
scope using the algorithm specified by property
\code{chord.scope.kind} and then writes the list of methods deemed
reachable and the reflection resolved by that algorithm to the files
specified by properties \code{chord.methods.file} and
\code{chord.reflect.file}, respectively.

The possible legal values of property \code{chord.scope.kind} are
[\code{dynamic}$|$\code{rta}$|$\code{cha}].  In each case, \Chord\ at
least expects properties \code{chord.main.class} and
\code{chord.class.path} to be set.

\begin{itemize}
\item
The {\tt dynamic} value instructs \Chord\ to compute program scope
dynamically, by running the program and observing using JVMTI the
classes that are loaded at run-time.  The number of times the program
is run and the command-line arguments to be supplied to the program in
each run is specified by properties \code{chord.run.ids} and
\code{chord.args.<id>} for each run ID {\tt <id>}.  By default, the
program is run only once, using run ID {\tt 0}, and without any
command-line arguments.  Only classes loaded in some run are regarded
as reachable but {\it all} methods of each loaded class are regarded
as reachable regardless of whether they were invoked in the run.  The
rationale behind this decision is to both reduce run-time overhead of
JVMTI and to increase the predictive power of program analyses
performed using the computed program scope.

\item
The {\tt rta} value instructs \Chord\ to compute program scope
statically using Rapid Type Analysis (RTA).

RTA is an iterative fixed-point algorithm.  It maintains a set of
reachable methods $M$.  The initial iteration starts by assuming that
only the main method in the main class is reachable (\Chord\ also
handles class initializer methods but we ignore them here for brevity;
we also ignore the set of reachable classes maintained besides the set
of reachable methods).  All object allocation sites $H$ contained in
methods in $M$ are deemed reachable (i.e., control-flow within method
bodies is ignored).  Whenever a dynamically-dispatching method call
site (i.e., an invokevirtual or invokeinterface site) with receiver of
static type $t$ is encountered in a method in $M$, only subtypes of
$t$ whose objects are allocated at some site in $H$ are considered to
determine the possible target methods, and each such target method is
added to $M$.  The process terminates when no more methods can be
added.

%RTA is a relatively inexpensive and precise algorithm in practice.  Its key shortcoming is that it makes
%no attempt to resolve reflection, which is rampant in real-world Java programs, and can therefore be
%unsound (i.e., underestimate the set of reachable classes and methods).
%The next option attempts to overcome this problem.
%
%\item
%The \code{rta_reflect} value instructs \Chord\ to compute program scope statically using Rapid Type Analysis
%and, moreover, to resolve a common reflection pattern:
%
%\begin{quote}
%\begin{verbatim}
%String s = ...;
%Class c = Class.forName(s);
%Object o = c.newInstance();
%T t = (T) o;
%\end{verbatim}
%\end{quote}
%
%This analysis is identical to RTA except that it additionally inspects every cast statement in the program,
%such as the last statement in the above snippet,
%and queries the class hierarchy to find all concrete classes that subclass \code{T} (if \code{T} is a class)
%or that implement \code{T} (if \code{T} is an interface).
%\Chord\ allows users to control which classes are included in the class hierarchy (see below).
%
\item
The {\tt cha} value instructs \Chord\ to compute program scope
statically using Class Hierarchy Analysis (CHA).

The key difference between CHA and RTA is that for invokevirtual and
invokeinterface sites with receiver of static type $t$, CHA considers
{\it all} subtypes of $t$ in the class hierarchy to determine the
possible target methods, whereas RTA restricts them to types of
objects allocated in methods deemed reachable so far.  As a result,
CHA is highly imprecise in practice, and also expensive since it
grossly overestimates the set of reachable classes and methods.
Nevertheless, \Chord\ allows users to control which classes are
included in the class hierarchy (see below) and thereby control the
precision and cost of CHA.

@@

The class hierarchy is built if property \code{chord.scope.kind} has
value {\tt cha}.  Users can control which classes are included in
building the class hierarchy by setting property \code{chord.ch.kind},
whose possible legal values are [\code{static}$|$\code{dynamic}].
\Chord\ first constructs the entire classpath of the given program by
concatenating in order the following classpaths:

\begin{enumerate}
\item
The boot classpath, specified by property \code{sun.boot.class.path}.
\item
The library extensions classpath, comprising all jar files in
directory \code{[java.home]/lib/dir/}.
\item
\Chord's classpath, specified by property
\code{chord.main.class.path}.  This includes the classpath of any
user-defined Java analyses, specified by user-defined property
\code{chord.java.analysis.path}, which is empty by default.
\item
The classpath of the given program, specified by user-defined property
\code{chord.class.path}, which is empty by default.
\end{enumerate}

All classes in the entire classpath (resulting from items 1--4 above)
are included in the class hierarchy with the following exceptions:
\begin{itemize}
\item
Duplicate classes, i.e., classes with the same name occurring in more
than one classpath element; in this case, all occurrences except the
first are excluded.
\item
Any class whose name's prefix is specified in the value of property
\code{chord.scope.exclude}.
\item
If property \code{chord.ch.kind} has value {\tt dynamic}, then
\Chord\ runs the given program and observes the set of all classes the
JVM loads; any class not in this set is excluded.  By default,
property \code{chord.ch.kind} has value {\tt static}.
\item
If the superclass of a class C is missing or if an interface
implemented/extended by a class/interface C is missing, where
``missing" means that it is either not in the classpath resulting from
items 1--4 above or it is excluded by one of these rules, then C
itself is excluded.  Note that this rule is recursive, e.g., if C has
superclass B which in turn has superclass A, and A is missing, then
both B and C are excluded.
\end{itemize}

>>>>

Chord computes the scope (i.e., reachable methods) of the given
program either if property \code{chord.build.scope} is set to {\tt
  true} or if some other task (e.g., a program analysis specified via
property \code{chord.run.analyses}) demands it by calling method
\code{chord.program.Program.g()}.

@@

If property \code{chord.reuse.scope} has value {\tt true} and both
files specified by properties \code{chord.methods.file} and
\code{chord.reflect.file} exist, then Chord regards those files as
specifying which methods to consider reachable and how to resolve
reflection, respectively.

@@

If property \code{chord.reuse.scope} has value {\tt false} or the
files specified by properties \code{chord.methods.file} or
\code{chord.reflect.file} do not exist, then Chord computes program
scope using the algorithm specified by property
\code{chord.scope.kind} and then writes the list of methods deemed
reachable and the reflection resolved by that algorithm to the files
specified by properties \code{chord.methods.file} and
\code{chord.reflect.file}, respectively.

The possible legal values of property \code{chord.scope.kind} are
[\code{dynamic}$|$\code{rta}$|$\code{cha}].  In each case, Chord at
least expects properties \code{chord.main.class} and
\code{chord.class.path} to be set.

\begin{itemize}
\item
The {\tt dynamic} value instructs Chord to compute program scope
dynamically, by running the program and observing using JVMTI the
classes that are loaded at run-time.  The number of times the program
is run and the command-line arguments to be supplied to the program in
each run is specified by properties \code{chord.run.ids} and
\code{chord.args.<id>} for each run ID {\tt <id>}.  By default, the
program is run only once, using run ID {\tt 0}, and without any
command-line arguments.  Only classes loaded in some run are regarded
as reachable but {\it all} methods of each loaded class are regarded
as reachable regardless of whether they were invoked in the run.  The
rationale behind this decision is to both reduce run-time overhead of
JVMTI and to increase the predictive power of program analyses
performed using the computed program scope.

\item
The {\tt rta} value instructs Chord to compute program scope
statically using Rapid Type Analysis (RTA).

RTA is an iterative fixed-point algorithm.  It maintains a set of
reachable methods $M$.  The initial iteration starts by assuming that
only the main method in the main class is reachable (Chord also
handles class initializer methods but we ignore them here for brevity;
we also ignore the set of reachable classes maintained besides the set
of reachable methods).  All object allocation sites $H$ contained in
methods in $M$ are deemed reachable (i.e., control-flow within method
bodies is ignored).  Whenever a dynamically-dispatching method call
site (i.e., an invokevirtual or invokeinterface site) with receiver of
static type $t$ is encountered in a method in $M$, only subtypes of
$t$ whose objects are allocated at some site in $H$ are considered to
determine the possible target methods, and each such target method is
added to $M$.  The process terminates when no more methods can be
added.

%RTA is a relatively inexpensive and precise algorithm in practice.  Its key shortcoming is that it makes
%no attempt to resolve reflection, which is rampant in real-world Java programs, and can therefore be
%unsound (i.e., underestimate the set of reachable classes and methods).
%The next option attempts to overcome this problem.
%
%\item
%The \code{rta_reflect} value instructs Chord to compute program scope statically using Rapid Type Analysis
%and, moreover, to resolve a common reflection pattern:
%
%\begin{quote}
%\begin{verbatim}
%String s = ...;
%Class c = Class.forName(s);
%Object o = c.newInstance();
%T t = (T) o;
%\end{verbatim}
%\end{quote}
%
%This analysis is identical to RTA except that it additionally inspects every cast statement in the program,
%such as the last statement in the above snippet,
%and queries the class hierarchy to find all concrete classes that subclass \code{T} (if \code{T} is a class)
%or that implement \code{T} (if \code{T} is an interface).
%Chord allows users to control which classes are included in the class hierarchy (see below).
%
\item
The {\tt cha} value instructs Chord to compute program scope
statically using Class Hierarchy Analysis (CHA).

The key difference between CHA and RTA is that for invokevirtual and
invokeinterface sites with receiver of static type $t$, CHA considers
{\it all} subtypes of $t$ in the class hierarchy to determine the
possible target methods, whereas RTA restricts them to types of
objects allocated in methods deemed reachable so far.  As a result,
CHA is highly imprecise in practice, and also expensive since it
grossly overestimates the set of reachable classes and methods.
Nevertheless, Chord allows users to control which classes are
included in the class hierarchy (see below) and thereby control the
precision and cost of CHA.

@@

The class hierarchy is built if property \code{chord.scope.kind} has
value {\tt cha}.  Users can control which classes are included in
building the class hierarchy by setting property \code{chord.ch.kind},
whose possible legal values are [\code{static}$|$\code{dynamic}].
Chord first constructs the entire classpath of the given program by
concatenating in order the following classpaths:

\begin{enumerate}
\item
The boot classpath, specified by property \code{sun.boot.class.path}.
\item
The library extensions classpath, comprising all jar files in
directory \code{[java.home]/lib/dir/}.
\item
Chord's classpath, specified by property
\code{chord.main.class.path}.  This includes the classpath of any
user-defined Java analyses, specified by user-defined property
\code{chord.java.analysis.path}, which is empty by default.
\item
The classpath of the given program, specified by user-defined property
\code{chord.class.path}, which is empty by default.
\end{enumerate}

All classes in the entire classpath (resulting from items 1--4 above)
are included in the class hierarchy with the following exceptions:
\begin{itemize}
\item
Duplicate classes, i.e., classes with the same name occurring in more
than one classpath element; in this case, all occurrences except the
first are excluded.
\item
Any class whose name's prefix is specified in the value of property
\code{chord.scope.exclude}.
\item
If property \code{chord.ch.kind} has value {\tt dynamic}, then
Chord runs the given program and observes the set of all classes the
JVM loads; any class not in this set is excluded.  By default,
property \code{chord.ch.kind} has value {\tt static}.
\item
If the superclass of a class C is missing or if an interface
implemented/extended by a class/interface C is missing, where
``missing" means that it is either not in the classpath resulting from
items 1--4 above or it is excluded by one of these rules, then C
itself is excluded.  Note that this rule is recursive, e.g., if C has
superclass B which in turn has superclass A, and A is missing, then
both B and C are excluded.
\end{itemize}

### 2010-12-13 main/doc/download_and_installation.tex
minor (added documentation + fixed javadoc tags)
migrated documentation from latex2html to hyperlatex

\begin{itemize}
\item
JDK 6 or higher, e.g. from \htmladdnormallink{IBM}{http://www.ibm.com/developerworks/java/jdk/} or
\htmladdnormallink{Sun}{http://java.sun.com/javase/downloads/index.jsp}
\item
\htmladdnormallink{Apache Ant}{http://ant.apache.org/}, a Java-based build tool
\item
a C++ compiler, e.g. \htmladdnormallink{GCC}{http://gcc.gnu.org/}
\item
a Make utility, e.g. \htmladdnormallink{GNU Make}{http://www.gnu.org/software/make/}
\item
\htmladdnormallink{Cygwin}{http://www.cygwin.com/}, if it is a Windows machine
\end{itemize}

Download a \Chord\ source release from \url{http://code.google.com/p/jchord/downloads/list} or
the latest source from the SVN repository at \url{http://code.google.com/p/jchord/source/checkout}.

@@

\begin{quote}
\begin{verbatim}
prompt> ant
\end{verbatim}
\end{quote}

\noindent To compile \Chord, run the following command in the same directory:

\begin{quote}
\begin{verbatim}
prompt> ant compile
\end{verbatim}
\end{quote}

\noindent This will compile the following:
\begin{itemize}
\item
the Java source code of \Chord\ from \code{main/src/java/} (and from \code{extra/src/java/} if that directory is present) to Java bytecode in \code{main/classes/}.
\item
the C source code of BDD library \htmladdnormallink{BuDDy}{http://buddy.sourceforge.net/} from \code{main/src/bdd/} to a shared library in \code{main/lib/} (\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and \code{libbuddy.dylib} on MacOS); this library is needed for executing program analyses written in Datalog using \htmladdnormallink{bddbddb}{http://bddbddb.sourceforge.net/} (a BDD-based Datalog solver).
\item
the C++ source code of the Chord instrumentation agent from \code{main/src/agent/} to a shared library in \code{main/lib/} (\code{libchord_instr_agent.so} on all architectures); this agent is needed for computing program scope dynamically and for executing dynamic program analyses.
\end{itemize}

>>>>

\begin{itemize}
\item
JDK 5 or higher, e.g. from
\xlink{IBM}{http://www.ibm.com/developerworks/java/jdk/} or
\xlink{Sun}{http://java.sun.com/javase/downloads/index.jsp}
\item
\xlink{Apache Ant}{http://ant.apache.org/}, a Java-based build tool
\item
a C++ compiler, e.g. \xlink{GCC}{http://gcc.gnu.org/}
\item
a Make utility, e.g. \xlink{GNU Make}{http://www.gnu.org/software/make/}
\item
\xlink{Cygwin}{http://www.cygwin.com/}, if it is a Windows machine
\end{itemize}

Download the latest Chord source code from the SVN repository:

\begin{verbatim}
    prompt> svn checkout http://jchord.googlecode.com/svn/trunk/ chord
\end{verbatim}
For many users, it may be sufficient to only checkout directory {\tt
  trunk/main/} instead of the entire {\tt trunk/}.  Other directories
of interest are {\tt trunk/test/}, which contains regression tests,
and {\tt trunk/libsrc/}, which contains the source code of software
used by Chord (Joeq, Javassist, bddbddb, JavaBDD, and jwutil).  All
other directories under {\tt trunk/} are projects built atop Chord
that might be of interest to some users.

@@

\begin{verbatim}
    prompt> ant
\end{verbatim}

\noindent To compile Chord, run the following command in the same
directory:

\begin{verbatim}
    prompt> ant compile
\end{verbatim}

\noindent This will compile the following:
\begin{itemize}
\item
the Java source code of Chord from \code{main/src/} to Java bytecode
in \code{main/classes/}.
\item
the C source code of BDD library
\xlink{BuDDy}{http://buddy.sourceforge.net/} from
\code{main/bdd/} to a shared library in \code{main/lib/}
(\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
\code{libbuddy.dylib} on MacOS); this library is needed for executing
program analyses written in Datalog using
\xlink{bddbddb}{http://bddbddb.sourceforge.net/} (a
BDD-based Datalog solver).
\item
the C++ source code of the Chord instrumentation agent from
\code{main/agent/} to a shared library in \code{main/lib/}
(\code{libchord_instr_agent.so} on all architectures); this agent is
needed for computing program scope dynamically and for executing
dynamic program analyses.
\end{itemize}

### 2010-12-13 main/doc/tuning_datalog_analysis.tex
migrated documentation from latex2html to hyperlatex

There are several tricks analysis writers can try to improve the
performance of \bddbddb, the Datalog solver used by \Chord, often
by several orders of magnitude:

\begin{enumerate}
\item
Set system properties \verb+noisy=yes+, \verb+tracesolve=yes+, and \verb+fulltracesolve=yes+
while running \bddbddb\ and observe which rule gets ``stuck" (i.e., takes several seconds to solve).
\verb+fulltracesolve+ is seldom useful, but \verb+noisy+ and \verb+tracesolve+ is
often very useful.  Once you identify the rule that is getting stuck, it
will also tell you which relations and which domains used in that rule
and which operation on them is taking a long time to solve.  Then try
to fix the slowdown problem with that rule by either simplifying the
relations involved (e.g., breaking their set of attributes into two if
possible) or changing the relative order of the domains of those
relations in the BDD variable order (doing the latter alone
frequently causes the problem to go away).

\item
Once you have ensured that none of the rules is getting ``stuck",
you will notice that some rules are applied too many times, and so
although each application of the rule itself isn't taking too much
time, the cumulative time for the rule is too much.  After finishing
solving a Datalog analysis, \bddbddb\ prints how long each rule took to
solve (both in terms of the number of times it was applied and the
cumulative time it took).  It sorts the rules in the order of the
cumulative time.  You need to focus on the rules that took the most
time to solve (they will be at the bottom of the list).  Assuming you
removed the problem of rules getting ``stuck", the rules will roughly
be in the order of the number of times they were applied.  Here is an
example:

@@

Notice that the second rule was applied 5031 times whereas the first
was applied 2871 times.  More importantly, the second rule took 6972
milliseconds in all, compared to 6798 for the first rule.  Hence, you
should focus on the second rule first, and try to speed it up.  This
means that you should focus only on relations IM, reachableI, specIMV,
and VH, and the domains I0, M0, V0, and H0.  Any changes you make that
do not affect these relations and domains are unlikely to make your
solving faster.  In general, look at the last few rules, not just the
last one, and try to identify the "sub-program" of the Datalog program
that seems problematic, and then focus on speeding up just that sub-
program.

\item
Tweak the BDD variable ordering in the Datalog file (the line that
starts with \verb+bddvarorder+).  This is one of the most
effective ways of making \bddbddb\ run faster.  As explained above, use
\verb+noisy+ and \verb+tracesolve+ to find out which pairs of domains in the
bddvarorder are responsible for the slowdown, and try to change their
relative ordering (note that you can use either `\_' or `x' between a pair
of domains, and the latter is commutative).

\item
You can add the \verb+.split+ keyword at the end of certain rules as a
hint to \bddbddb\ to break up those rules into simpler ones that can be
solved faster.  You can also set property \verb+split_all_rules=yes+ as shorthand
for splitting all rules without adding the \verb+.split+ keyword to any of
them, though I seldom find splitting all rules helpful.

\item
You can yourself break down rules by creating intermediate relations (the more
relations you have on the RHS of a rule the slower it takes to solve
that rule).  This is another very effective way to make \bddbddb\ run
faster.

\item
Try breaking down a single Datalog program into two programs.  Of
course, you cannot separate mutually-recursive rules into two
different programs, but if you unnecessarily club
together rules that could have gone into different programs, then they
can put conflicting demands on \bddbddb\ (e.g., on \verb+bddvarorder+).
So if rule 2 uses the result of rule 1 and rule 1 does not use the result of
rule 2, then put rule 1 and rule 2 in two different Datalog programs.

\item
Observe the sizes of the BDDs representing the relations that are
input and output.  \bddbddb\ tells both the number of tuples in each
relation and the number of "nodes" in the BDD.  Try changing the
bddvarorder for the domains of the relation, and observe how the
number of "nodes" in the bdd for that relation change.  You will
notice that some orderings do remarkably better than others.  Then
note down these orderings as invariants that you will not violate as
you tweak other things.

\item
The relative ordering of values *within* domains (e.g.
in domains named \verb+M+, \verb+H+, \verb+C+, etc. in Chord) affects the
solving time of \bddbddb, but
I've never tried changing this and studying its effect.  It might be
worth trying.  For instance, John Whaley's PLDI'04 paper describes a
specific way in which he numbers contexts (in domain \verb+C+) and that it was
fundamental to the speedup of his infinity-CFA points-to analysis.

\item
Finally, it is worth emphasizing that BDDs are not magic.
If your algorithm itself is fundamentally hard to scale, then BDDs are
unlikely to help you a whole lot.  Secondly, many things are awkward to
encode as integers (e.g., the abstract contexts in the domain \verb+C+ 
in Chord) or as Datalog rules.
For instance, I've noticed that summary-based context-sensitive program
analyses are hard to express in Datalog.  The may-happen-in-parallel
analysis provided in Chord shows a relatively simple kind of summary-based
analysis that uses the Reps-Horwitz-Sagiv tabulation algorithm.  But this
is as far as I could get---more complicated summary-based algorithms are
best written in Java itself instead of Datalog.
\end{enumerate}

>>>>

There are several tricks analysis writers can try to improve the
performance of bddbddb, the Datalog solver used by Chord, often
by several orders of magnitude.
Try these tricks by running the following command:

\begin{verbatim}
    prompt> ant -Ddlog.file=<file> -Dwork.dir=<dir> solve
\end{verbatim}
where {\tt <file>} denotes the file defining the Datalog analysis
to be tuned, and {\tt <dir>} is the directory containing the
program domains ({\tt *.dom} files) and program relations ({\tt *.bdd} files)
consumed by the analysis (this is by default the
\code{chord_output/bddbddb/} directory generated
by a previous run of Chord.

\begin{enumerate}
\item
Set properties \verb+noisy=yes+, \verb+tracesolve=yes+, and \verb+fulltracesolve=yes+
on the above command line and observe which rule gets ``stuck" (i.e., takes several seconds to solve).
\verb+fulltracesolve+ is seldom useful, but \verb+noisy+ and \verb+tracesolve+ are
often very useful.  Once you identify the rule that is getting stuck, it
will also tell you which relations and which domains used in that rule,
and which operation on them, is taking a long time to solve.  Then try
to fix the problem with that rule by doing either or both of the following:
\begin{itemize}
\item
Break down the rule into multiple rules by creating intermediate relations (the more
relations you have on the RHS of a rule the slower it generally takes to solve
that rule).
\item
Change the relative order of the domains of those
relations in the BDD variable order
(note that you can use either `\_' or `x' between a pair of domains).
\end{itemize}

\item
Once you have ensured that none of the rules is getting ``stuck",
you will notice that some rules are applied too many times, and so
although each application of the rule itself isn't taking too much
time, the cumulative time for the rule is too much.  After finishing
solving a Datalog analysis, bddbddb prints how long each rule took to
solve (both in terms of the number of times it was applied and the
cumulative time it took).  It sorts the rules in the order of the
cumulative time.  You need to focus on the rules that took the most
time to solve (they will be at the bottom of the list).  Assuming you
removed the problem of rules getting ``stuck", the rules will roughly
be in the order of the number of times they were applied.  Here is an
example:

@@

Notice that the second rule was applied 5031 times whereas the first
was applied 2871 times.  More importantly, the second rule took 6972
milliseconds in all, compared to 6798 for the first rule.  Hence, you
should focus on the second rule first, and try to speed it up.  This
means that you should focus only on relations IM, reachableI, specIMV,
and VH, and the domains I0, M0, V0, and H0.  Any changes you make that
do not affect these relations and domains are unlikely to make your
solving faster.  In general, look at the last few rules, not just the
last one, and try to identify the ``sub-analysis" of the Datalog analysis
that seems problematic, and then focus on speeding up just that sub-
analysis.

\item
You can add the \verb+.split+ keyword at the end of certain rules as a
hint to bddbddb to decompose those rules into simpler ones that can be
solved faster.  You can also set property \verb+split_all_rules=yes+ as shorthand
for splitting all rules without adding the \verb+.split+ keyword to any of
them, though I seldom find splitting all rules helpful.

\item
You can try to decompose a single Datalog analysis file into two separate Datalog analysis
files.  Of course, you cannot separate mutually-recursive rules into two
different analyses, but if you unnecessarily club
together rules that could have gone into different analyses, then they
can put conflicting demands on bddbddb (e.g., on the BDD variable order).
So if rule 2 uses the result of rule 1 and rule 1 does not use the result of
rule 2, then put rule 1 and rule 2 in separate Datalog analyses.

\item
Observe the sizes of the BDDs representing the relations that are
input and output.  bddbddb prints both the number of tuples in each
relation and the number of nodes in the BDD.  Try changing the
BDD variable order for the domains of the relation, and observe how the
number of nodes in the BDD for that relation change.  You will
notice that some orders perform remarkably better than others.  Then
note down these orders as invariants that you will not violate as
you tweak other things.

\item
The relative order of values *within* domains (e.g.,
in domains named \verb+M+, \verb+H+, \verb+C+, etc. in Chord) affects the
performance of bddbddb, but
I've never tried changing this and studying its effect.  It might be
worth trying.  For instance, John Whaley's PLDI'04 paper describes a
specific way in which he numbers contexts (in domain \verb+C+) and that it was
fundamental to the speedup of his ``infinity"-CFA points-to analysis.

\item
Finally, it is worth emphasizing that BDDs are not magic.
If your algorithm itself is fundamentally hard to scale, then BDDs are
unlikely to help you a whole lot.  Secondly, many things are awkward to
encode as integers (e.g., the abstract contexts in domain \verb+C+ 
in Chord) or as Datalog rules.
For instance, I've noticed that summary-based context-sensitive program
analyses are hard to express in Datalog.  The may-happen-in-parallel
analysis provided in Chord shows a relatively simple kind of summary-based
analysis that uses the Reps-Horwitz-Sagiv tabulation algorithm.  But this
is as far as I could get---more complicated summary-based algorithms are
best written in Java itself instead of Datalog.
\end{enumerate}

### 2010-12-13 main/doc/whatis_chord.tex
migrated documentation from latex2html to hyperlatex

\Chord\ is a program analysis platform for Java designed with the
primary goal of helping advance the state-of-the-art in the field by
enabling researchers to productively design, implement, and evaluate a
broad variety of program analyses. To this end, Chord has the
following key characteristics:

@@

\noindent \Chord\ is intended to work on a variety of platforms,
including Linux, Windows/Cygwin, and MacOS.  It has been tested most
extensively on Linux.  It is open source software distributed under
the \htmladdnormallink{New BSD
  License}{http://www.opensource.org/licenses/bsd-license.php}.
Improvements by users are welcome and encouraged.  The project website
is \url{http://berkeley.intel-research.net/chord/} and the software
website is \url{http://code.google.com/p/jchord/}.

>>>>

Chord is a program analysis platform for Java designed with the
primary goal of helping advance the state-of-the-art in the field by
enabling researchers to productively design, implement, and evaluate a
broad variety of program analyses. To this end, Chord has the
following key characteristics:

@@

\noindent Chord is intended to work on a variety of platforms,
including Linux, Windows/Cygwin, and MacOS.  It has been tested most
extensively on Linux.  It is open source software distributed under
the \xlink{New BSD License}{http://www.opensource.org/licenses/bsd-license.php}.
Improvements by users are welcome and encouraged.  The project website
is \url{http://berkeley.intel-research.net/chord/} and the software
website is \url{http://code.google.com/p/jchord/}.

### 2010-12-13 main/doc/writing_dynamic_analysis.tex
migrated documentation from latex2html to hyperlatex

You can change the default values of various properties for configuring your dynamic analysis;
see Sections \ref{sec:scope-props} and \ref{sec:instr-props}.
For instance:
\begin{itemize}
\item
You can set property \code{chord.scope.kind} to {\tt dynamic} so that the program scope
is computed dynamically (i.e., by running the program) instead of statically.
\item
You can exclude certain classes (e.g., JDK classes) from being instrumented by setting
properties \code{chord.std.scope.exclude}, \code{chord.ext.scope.exclude}, and
\code{chord.scope.exclude}.
\item
You can choose between online (i.e. load-time) and offline bytecode instrumentation by
setting property \code{chord.instr.kind} to {\tt online} or {\tt offline}.
\item
You can require the event-generating and event-handling JVMs to be one and the same (by setting
property \code{chord.trace.kind} to {\tt none}),
or to be separate (by setting property {\tt chord.trace.kind} to {\tt full} or {\tt pipe}, depending upon
whether you want the two JVMs to exchange events by a regular file or a POSIX pipe, respectively).
Using a single JVM can cause correctness/performance issues if event-handling Java code itself is instrumented
(e.g., say the event-handling code uses class \code{java.util.ArrayList} which is not excluded from program scope).
Using separate JVMs prevents such issues since the event-handling JVM runs uninstrumented bytecode (only
the event-generating JVM runs instrumented bytecode).
If a regular file is used to exchange events between the two JVMs, then the JVMs run serially:
the event-generating JVM first runs to completion, dumps the entire dynamic trace to the regular file,
and then the event-handling JVM processes the dynamic trace.
If a POSIX pipe is used to exchange events between the two JVMs, then the JVMs run in lockstep.
Obviously, a pipe is more efficient for very long traces, but it 
not portable (e.g., it does not currently work on Windows/Cygwin), and the traces cannot be reused
across Chord runs (see the following item).
\item
You can reuse dynamic traces from a previous \Chord\ run by setting property
\code{chord.reuse.traces} to {\tt true}.  In this case, you must also set property
{\tt chord.trace.kind} to {\tt full}.
\item
You can set property \code{chord.dynamic.haltonerr} to {\tt false} to prevent \Chord\ from terminating
even if the program on which dynamic analysis is being performed crashes.
\end{itemize}

\Chord\ offers much more flexibility in crafting dynamic analyses.
You can define your own instrumentor (by subclassing \code{chord.instr.CoreInstrumentor}
instead of using the default \code{chord.instr.Instrumentor}) and your own event handler (by subclassing \code{chord.runtime.CoreEventHandler}
instead of using the default \code{chord.runtime.EventHandler}).
You can ask the dynamic analysis to use your custom instrumentor and/or your custom event handler by overriding
methods \code{getInstrumentor()} and \code{getEventHandler()}, respectively, defined in \code{chord.project.analyses.CoreDynamicAnalysis}.
Finally, you can define your own dynamic analysis template by subclassing \code{chord.project.analyses.CoreDynamicAnalysis}
instead of subclassing the default \code{chord.project.analyses.DynamicAnalysis}.

>>>>

You can change the default values of various properties for configuring your dynamic analysis;
see Sections \ref{sec:scope-props} and \ref{sec:instr-props}.
For instance:
\begin{itemize}
\item
You can set property \code{chord.scope.kind} to {\tt dynamic} so that the program scope
is computed dynamically (i.e., by running the program) instead of statically.
\item
You can exclude certain classes (e.g., JDK classes) from being instrumented by setting
properties \code{chord.std.scope.exclude}, \code{chord.ext.scope.exclude}, and
\code{chord.scope.exclude}.
\item
You can choose between online (i.e. load-time) and offline bytecode instrumentation by
setting property \code{chord.instr.kind} to {\tt online} or {\tt offline}.
\item
You can require the event-generating and event-handling JVMs to be one and the same (by setting
property \code{chord.trace.kind} to {\tt none}),
or to be separate (by setting property {\tt chord.trace.kind} to {\tt full} or {\tt pipe}, depending upon
whether you want the two JVMs to exchange events by a regular file or a POSIX pipe, respectively).
Using a single JVM can cause correctness/performance issues if event-handling Java code itself is instrumented
(e.g., say the event-handling code uses class \code{java.util.ArrayList} which is not excluded from program scope).
Using separate JVMs prevents such issues since the event-handling JVM runs uninstrumented bytecode (only
the event-generating JVM runs instrumented bytecode).
If a regular file is used to exchange events between the two JVMs, then the JVMs run serially:
the event-generating JVM first runs to completion, dumps the entire dynamic trace to the regular file,
and then the event-handling JVM processes the dynamic trace.
If a POSIX pipe is used to exchange events between the two JVMs, then the JVMs run in lockstep.
Obviously, a pipe is more efficient for very long traces, but it 
not portable (e.g., it does not currently work on Windows/Cygwin), and the traces cannot be reused
across Chord runs (see the following item).
\item
You can reuse dynamic traces from a previous Chord run by setting property
\code{chord.reuse.traces} to {\tt true}.  In this case, you must also set property
{\tt chord.trace.kind} to {\tt full}.
\item
You can set property \code{chord.dynamic.haltonerr} to {\tt false} to prevent Chord from terminating
even if the program on which dynamic analysis is being performed crashes.
\end{itemize}

Chord offers much more flexibility in crafting dynamic analyses.
You can define your own instrumentor (by subclassing \code{chord.instr.CoreInstrumentor}
instead of using the default \code{chord.instr.Instrumentor}) and your own event handler (by subclassing \code{chord.runtime.CoreEventHandler}
instead of using the default \code{chord.runtime.EventHandler}).
You can ask the dynamic analysis to use your custom instrumentor and/or your custom event handler by overriding
methods \code{getInstrumentor()} and \code{getEventHandler()}, respectively, defined in \code{chord.project.analyses.CoreDynamicAnalysis}.
Finally, you can define your own dynamic analysis template by subclassing \code{chord.project.analyses.CoreDynamicAnalysis}
instead of subclassing the default \code{chord.project.analyses.DynamicAnalysis}.
